---
title: "Project Report for Basic Inferential Data Analysis"
author: "Yukai Zou"
output: html_document
---
```{r setup, include=FALSE}
opts_chunk$set(dev = 'pdf')
```

# Overview
This report is the second part of the Coursera statistical inference course project. In this report, the ToothGrowth data in the R datasets will be analyzed using some basic inferential statistics. Exploratory data analyses and summary of the data will be first given. After that, comparisons of tooth growth by supp and dose will be performed respectively, with the relevant confidence intervals calculated; different hypotheses and assumptions will be formulated and tested using T statistics. Conclusions will be made based on the statistical results.

# Exploratory Analyses on ToothGrowth Data
For exploratory plots, see Appendix for reference. Based on the plots, in terms of supplements (Figure 1), OJ seems to contribute to a higher tooth growth compared with VC; in terms of dose (Figure 2), a higher dose seems to increase the length of tooth growth. Both observations require hypothesis testing.

# Basic Summary of the Data
```{r, echo=FALSE}
library(datasets)
summary(ToothGrowth)
```

## Basic Summary of Tooth Growth by Supp
## Basic Summary of Tooth Growth by Dose

# Comparisons of Tooth Growth by Supp

Assuming that the data of tooth growth by OJ and VC are both normally distributed, and the population variances are unequal.

**H0**: there is no difference between tooth growth by VC and tooth growth by OJ (VC = OJ).

**Ha**: the tooth growth by VC is larger than the tooth growth by OJ (VC > OJ).

**Hypothesis testing**:
```{r, echo=FALSE}
library(datasets)
VC <- ToothGrowth$len[1:30]
OJ <- ToothGrowth$len[31:60]

mOJ <- mean(OJ); mVC <- mean(VC)
sdOJ <- sqrt(var(OJ)); sdVC <- sqrt(var(VC))
nOJ <- length(OJ); nVC <- length(VC)

df <- (sdOJ^2/nOJ + sdVC^2/nVC)^2/((sdOJ^2/30)^2/(30-1) + (sdVC^2/30)^2/(30-1))
CI <- mOJ - mVC + c(-1,1) * qt(0.975, df) * sqrt(sdOJ^2/nOJ + sdVC^2/nVC)
t.test(OJ, VC, paired = FALSE, var.equal = FALSE)
```

According to the statistical test, the confidence interval did not exclude zero, and the p-value is not less than the 0.05 threshold.

**Conclusion**: The null hypothesis is not rejected. There is no difference between tooth growth by VC and tooth growth by OJ.

# Comparisons of Tooth Growth by Dose
## Dose 0.5 v.s. Dose 1.0

Assuming that the data of tooth growth by dose 0.5 and dose 1.0 are normally distributed, and the population variances are equal.

**H0**: there is no difference between tooth growth by dose 0.5 and tooth growth by dose 1.0 (dose 0.5 = dose 1.0).

**Ha**: the tooth growth by dose 0.5 is smaller than the tooth growth by dose 1.0 (dose 0.5 < dose 1.0).

**Hypothesis testing**:
```{r, echo=FALSE}
library(datasets)
Dose12 <- subset(ToothGrowth, dose %in% c(0.5,1))
t.test(len ~ dose, paired = FALSE, var.equal = TRUE, data = Dose12)
```

According to the statistical test, both the lower and upper limit of the confidence interval is smaller than zero, and the p-value is much smaller than the 0.05 threshold.

**Conclusion**: The null hypothesis is rejected. The tooth growth by dose 0.5 is smaller than the tooth growth by dose 1.0.

## Dose 0.5 v.s. Dose 2.0

Assuming that the data of tooth growth by dose 0.5 and dose 2.0 are normally distributed, and the population variances are not equal.

**H0**: there is no difference between tooth growth by dose 0.5 and tooth growth by dose 2.0 (dose 0.5 = dose 2.0).

**Ha**: the tooth growth by dose 0.5 is smaller than the tooth growth by dose 2.0 (dose 0.5 < dose 2.0).

**Hypothesis testing**:
```{r, echo=FALSE}
library(datasets)
Dose13 <- subset(ToothGrowth, dose %in% c(0.5,2))
t.test(len ~ dose, paired = FALSE, var.equal = FALSE, data = Dose13)
```

According to the statistical test, both the lower and upper limit of the confidence interval is smaller than zero, and the p-value is much smaller than the 0.05 threshold.

**Conclusion**: The null hypothesis is rejected. The tooth growth by dose 0.5 is smaller than the tooth growth by dose 2.0.

## Dose 0.5 v.s. Dose 2.0

Assuming that the data of tooth growth by dose 1.0 and dose 2.0 are normally distributed, and the population variances are not equal.

**H0**: there is no difference between tooth growth by dose 1.0 and tooth growth by dose 2.0 (dose 1.0 = dose 1.0).

**Ha**: the tooth growth by dose 1.0 is smaller than the tooth growth by dose 2.0 (dose 1.0 < dose 2.0).

**Hypothesis testing**:
```{r, echo=FALSE}
library(datasets)
Dose23 <- subset(ToothGrowth, dose %in% c(1,2))
t.test(len ~ dose, paired = FALSE, var.equal = FALSE, data = Dose23)
```

According to the statistical test, both the lower and upper limit of the confidence interval is smaller than zero, and the p-value is much smaller than the 0.05 threshold.

**Conclusion**: The null hypothesis is rejected. The tooth growth by dose 1.0 is smaller than the tooth growth by dose 2.0.

# Appendix
## R Codes
To be added

## Exploratory Plots
```{r, echo=FALSE, fig.cap="Tooth Growth by Different Supp"}
library(datasets)
library(ggplot2)
qplot(supp, len, data = ToothGrowth, geom = "boxplot", fill = supp, main = "Tooth Growth by Different Supp")
```

```{r, echo=FALSE, fig.cap="Tooth Growth by Different Dose"}
library(datasets)
library(ggplot2)
len <- ToothGrowth$len
dose <- as.factor(ToothGrowth$dose)
g <- ggplot(data.frame(len, dose), aes(y = len,x = dose, fill = dose))
g <- g + geom_boxplot() + labs(title = "Tooth Growth by Different Dose", x = "dose", y = "len")
g
```
